---
title: "Economics Homework 2"
author: "Naomi Raal and Michelle Geldin"
date: "4/20/2022"
output: 
  html_document: 
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE, warning = FALSE)
library(here)
library(tidyverse)
library(janitor)
library(here)
```

```{r}
water <- read.csv(here("Water_Districts.csv")) 
```

## Question 1

```{r}
model.kern <- lm(MC_Kern ~ 0 + Reduction, data=water)
```

```{r}
B = model.kern$coefficient[1]
```

```{r}
predicted.mc <- B*seq(0, 150, 500)
```

```{r}
ggplot(data = water, aes(x = MC_Kern, y = Reduction)) +
  geom_point(color = "black") + geom_smooth(method = 'lm', se = FALSE) + labs(x = "Marginal Cost", y = "Reduction (Acre Feet)", title = "Kern Marginal Abatement Cost")
```


```{r}
model.mojave <- lm(MC_Mojave ~ 0 + Reduction, data=water)
```

```{r}
B = model.mojave$coefficient[1]
```

```{r}
predicted.mc <- B*seq(0, 150, 500)
```

```{r}
ggplot(data = water, aes(x = MC_Mojave, y = Reduction)) +
  geom_point(color = "black") + geom_smooth(method = 'lm', se = FALSE) + labs(x = "Marginal Cost", y = "Reduction (Acre Feet)", title = "Mojave Marginal Abatement Cost")
```


```{r}
model.antelope <- lm(MC_Antelope ~ 0 + Reduction, data=water)
```

```{r}
B = model.antelope$coefficient[1]
```

```{r}
predicted.mc <- B*seq(0, 150, 500)
```

```{r}
ggplot(data = water, aes(x = MC_Antelope, y = Reduction)) +
  geom_point(color = "black") + geom_smooth(method = 'lm', se = FALSE) + labs(x = "Marginal Cost", y = "Reduction (Acre Feet)", title = "Antelope Marginal Abatement Cost")
```


```{r}
model.ventura <- lm(MC_Ventura ~ 0 + Reduction, data=water)
```

```{r}
B = model.antelope$coefficient[1]
```

```{r}
predicted.mc <- B*seq(0, 150, 500)
```

```{r}
ggplot(data = water, aes(x = MC_Ventura, y = Reduction)) +
  geom_point(color = "black") + geom_smooth(method = 'lm', se = FALSE) + labs(x = "Marginal Cost", y = "Reduction (Acre Feet)", title = "Ventura Marginal Abatement Cost")
```

## Question 2



```{r}
model.kern <- lm(MC_Kern ~ 0 + Reduction, data = water)
summary(model.kern)

kern.baseline <- 150
kern_data <- tibble(Reduction = 0:150)
predicted_water_districts <- predict(model.kern, newdata = kern_data)
kern_data <- kern_data %>% 
  mutate(predicted_water_districts = predicted_water_districts)

```

```{r, include=FALSE}
summary(water)
```


```{r}
# define Kern demand curve
kern_data <- kern_data %>% 
  mutate(water_demand = kern.baseline-Reduction)
#plot demand curve
ggplot(kern_data, aes(x = water_demand, y = predicted_water_districts)) + 
  geom_point() + 
  geom_line() +
  labs(x = 'Water Demand', y = '$/AF', title = "Kern demand curve")
```


```{r}
# Define Mojave Demand Curve
model.mojave <- lm(MC_Mojave ~ 0 + Reduction, data = water)
summary(model.mojave)

mojave.baseline <- 140
mojave_data <- tibble(Reduction = 0:140)
predicted_water_districts <- predict(model.mojave, newdata = mojave_data)
mojave_data <- mojave_data %>% 
  mutate(predicted_water_districts = predicted_water_districts)
```


```{r}
mojave_data <- mojave_data %>% 
  mutate(water_demand = mojave.baseline-Reduction)
#plot demand curve
ggplot(mojave_data, aes(x = water_demand, y = predicted_water_districts)) + 
  geom_point() + 
  geom_line() +
  labs(x = 'Water Demand', y = '$/AF', title = "Mojave demand curve") 
```



```{r}
# Define Antelope Demand Curve
model.antelope <- lm(MC_Antelope ~ 0 + Reduction, data = water)
summary(model.antelope)

antelope.baseline <- 220
antelope_data <- tibble(Reduction = 0:220)
predicted_water_districts <- predict(model.antelope, newdata = antelope_data)
antelope_data <- antelope_data %>% 
  mutate(predicted_water_districts = predicted_water_districts)
```

```{r}
antelope_data <- antelope_data %>% 
  mutate(water_demand = antelope.baseline-Reduction)
#plot demand curve
ggplot(antelope_data, aes(x = water_demand, y = predicted_water_districts)) + 
  geom_point() + 
  geom_line() +
  labs(x = 'Water Demand', y = '$/AF', title = "Antelope demand curve")
```


```{r}
# Ventura Demand Curve 
model.ventura <- lm(MC_Ventura ~ 0 + Reduction, data = water)
summary(model.ventura)

ventura.baseline <- 245
ventura_data <- tibble(Reduction = 0:245)
predicted_water_districts <- predict(model.ventura, newdata = ventura_data)
ventura_data <- ventura_data %>% 
  mutate(predicted_water_districts = predicted_water_districts)
```

```{r}
ventura_data <- ventura_data %>% 
  mutate(water_demand = ventura.baseline-Reduction)
#plot demand curve
ggplot(ventura_data, aes(x = water_demand, y = predicted_water_districts)) + 
  geom_point() + 
  geom_line() +
  labs(x = 'Water Demand', y = '$/AF', title = "Ventura demand curve")
```

```{r}
## combining 
kern.baseline <- 150
kern_data <- tibble(Reduction = 0:150)
predicted_water_districts <- predict(model.kern, newdata = kern_data)
kern_data <- kern_data %>% mutate(predicted_water_districts = predicted_water_districts)

#use as baseline - abatement 

kern_data <- kern_data %>% mutate(water_demand = kern.baseline-Reduction)
#using the demand data for kern and mojave, join them into one dataset for plotting

joined_demand <- bind_rows('Mojave' = mojave_data,
                           'Kern' = kern_data, 'Antelope' = antelope_data, 'Ventura' = ventura_data, .id = 'Districts')
ggplot(data = joined_demand, aes(x = water_demand, y = predicted_water_districts, color = Districts)) + geom_line() + labs(title = "Each District's Demand Curve for Water")
```

Based on our graph and calculations, Antelope would be willing to pay the most for the first AF of water.

## Question 3

#### part a. (cap without trade)

```{r}
# Reduce each district’s water use by the same fraction (e.g., 1/3 of current baseline use), so the 500 AF target is met. (multiplier is 0.662)
fraction <- 0.662
kern_use_cap <- 150*fraction
#50.7 reduced 
mojave_use_cap <- 140*fraction 
#47.32 reduced
antelope_use_cap <- 220*fraction
#74.36 reduced
ventura_use_cap <- 245*fraction
#82.81 reduced
#mojave_cap+ventura_cap+antelope_cap+kern_cap --> This all adds up to 500 AF, the total use 
# But cost is a function of abatement, so we need the abatement values instead.
kern_abate_cap <- 150-kern_use_cap
mojave_abate_cap <- 140-mojave_use_cap
antelope_abate_cap <- 220-antelope_use_cap
ventura_abate_cap <- 245-ventura_use_cap
# DON'T KNOW ABOUT STRINGENCY HERE
# Cost, plug in REDUCTION to each district's cost curve, then take the area under the marginal cost curve for each at the given price/quantity reduction
kern_price_cap <- 2.29*kern_abate_cap
kern_cost_cap <- 0.5*kern_price_cap*kern_abate_cap
mojave_price_cap <- 3.8*mojave_abate_cap
mojave_cost_cap <- 0.5*mojave_price_cap*mojave_abate_cap
antelope_price_cap <- 2.86*antelope_abate_cap
antelope_cost_cap <- 0.5*antelope_price_cap*antelope_abate_cap
ventura_price_cap <- 1.78*ventura_abate_cap
ventura_cost_cap <- 0.5*ventura_price_cap*ventura_abate_cap
total_cost_cap = ventura_cost_cap + kern_cost_cap + antelope_cost_cap + mojave_cost_cap
```

*(1)* Each district has to reduce water use by `r (1-fraction)*100`%. 

*(2)* Total cost of this approach is $`r round(total_cost_cap,0)`.

*(3)* Individual district costs are:

- Kern : $`r round(kern_cost_cap,0)`
- Mojave: $`r round(mojave_cost_cap,0)`
- Antelope: $`r round(antelope_cost_cap,0)`
- Ventura: $`r round(ventura_cost_cap,0)`

*(4)* Cap without trade does not create tax revenue.

#### part b. (tax on water use) 

```{r}
# Tax applied to reach 500 AF level, 255 AF reduction 
# Calculate agg. demand to get the price at which cumulative water use is 500 AF
# Add together the inverse slopes of each MC curve 
aggregate_cost_slope <- (1/2.29 + 1/3.8 + 1/2.86 + 1/1.78)
tax_level <- 255/aggregate_cost_slope
# Tax of $158.26/AF is applied across all counties. 
# Now we need to see cost to each district, plug the price into the individual MC curves then take the area under each and sum for total. 
# Kern
# tax_level = 2.29A
kern_abate_tax <- tax_level/2.29
# 69.15 AF
kern_cost_tax <- 0.5*tax_level*kern_abate_tax
# Mojave
# tax_level = 3.8A
mojave_abate_tax <- tax_level/3.8
#41.67 AF
mojave_cost_tax <- 0.5*tax_level*mojave_abate_tax
#Antelope
# tax_level = 2.86A
antelope_abate_tax = tax_level/2.86
# 55.37 AF
antelope_cost_tax <- 0.5*tax_level*antelope_abate_tax
# Ventura
# tax_level = 1.78A 
ventura_abate_tax = tax_level/1.78
# 88.96
ventura_cost_tax <- 0.5*tax_level*ventura_abate_tax
#ventura_abate_tax+mojave_abate_tax+antelope_abate_tax+kern_abate_tax
# Total abatement still 255 
total_cost_tax = 0.5*255*tax_level 
# Revenue is the tax times the AF used. 
tax_revenue <- tax_level*500
```

*(1)* Stringency of this policy: $`r round(tax_level,0)`. 

*(2)* Total cost of this approach: $`r round(total_cost_tax,0)`.

*(3)* Individual district costs:

- Kern : $`r round(kern_cost_tax,0)`
- Mojave: $`r round(mojave_cost_tax,0)`
- Antelope: $`r round(antelope_cost_tax,0)`
- Ventura: $`r round(ventura_cost_tax,0)` 

*(4)* Tax revenue generated: $`r round(tax_revenue,0)`

#### part c. (cap and trade) 
From part (a), we calculated the cost to each district for the cap without trade to be: 

- Kern : $`r round(kern_cost_cap,0)`, reduction of 50.7 AF 
- Mojave: $`r round(mojave_cost_cap,0)`, reduction of 42.73 AF
- Antelope: $`r round(antelope_cost_cap,0)`, reduction of 74.36 AF
- Ventura: $`r round(ventura_cost_cap,0)`, reduction of 82.81 AF

We also calculated the total cost from above to be: $`r round(total_cost_cap,0)`.

```{r}
# Given the total marginal cost, we should assume these will be set equal for firms to get the most efficient/lowest cost solution. 
# THIS SHOULD BE THE SAME AS THE TAX IN TERMS OF ABATEMENT PER DISTRICT 
#Kern initially abates 50.7 AF at a price/MC of $115.6, they use 99.3 AF
#Mojave initially abates 42.73 AF at price/MC of $107.89, they use 92.7 AF
#Antelope initially abates 74.36 AF at price/MC of $169.54, they use 145.6 AF
#Ventura initially abates 82.8 AF at price/MC of $188.8, they use 162.2 AF
# Use post-quota could probably be considered their "allocations"/"rights"
# Still need to abate 255 AF total, but Mojave and Kern will likely face higher relative MC while Antelope and Ventura will face slightly lower MC. 
trade_mc <- tax_level
#Kern Q with trade
kern_abate_trade <- trade_mc/2.29
#69.15 AF is abatement level (far higher than initial)
#Mojave Q with trade
mojave_abate_trade <- trade_mc/3.8
#41.67 AF is new abatement level (almost same as initial)
#Antelope Q with trade
antelope_abate_trade <- trade_mc/2.86
#55.37 AF is new abatement level (far below initial)
#Ventura Q with trade
ventura_abate_trade <- trade_mc/1.78
#88.96 AF is new abatement level (higher than initial)
#ventura_abate_trade+antelope_abate_trade+mojave_abate_trade+kern_abate_trade
# Same allocation/cost as efficient tax, different revenue implications. MC for abatement is the same for each district at these allocations. 
#ventura_cost_cap/ventura_abate_cap
#ventura_cost_tax/ventura_abate_trade
# MC lower compared to cap
#antelope_cost_cap/antelope_abate_cap
#antelope_cost_tax/antelope_abate_trade
# MC Lower compared to cap
#kern_cost_cap/kern_abate_cap
#kern_cost_tax/kern_abate_trade
# MC higher compared to cap
#mojave_cost_cap/mojave_abate_cap
#mojave_cost_tax/mojave_abate_trade
# MC higher compared to cap
total_cost_trade<- 0.5*trade_mc*255
```

*(1)* Stringency of this policy is different for each district: 

- Kern: **0.44**
- Mojave: **0.26**
- Antelope: **0.35**
- Ventura: **0.56**

*(2)* Total cost: $`r round(total_cost_trade,0)` (same as tax).
This is $`r round(total_cost_cap - total_cost_trade, 0)` lower than the cost of the cap with no trade.

*(3)* Individual district costs:

- Kern : $`r round(kern_cost_tax - trade_mc*18.47,0)`
- Mojave: $`r round(mojave_cost_tax - trade_mc*-5.64,0)`
- Antelope: $`r round(antelope_cost_tax - trade_mc*-18.98,0)`
- Ventura: $`r round(ventura_cost_tax - trade_mc*6.17,0)` 

*(4)* Tax revenue is not generated from this policy. The thought behind this strategy is that districts with lower costs (Mojave, Antelope) would pay districts with higher costs (Kern, Ventura) to incentive them to abate more than they would otherwise.


### Question 4

#### Drought and cap without trade

```{r}
# Cap without trade. Reduce each district’s water use by the same fraction, so the 300 AF target is met. (multiplier is 0.397)
fraction_drought <- 0.397
kern_use_d <- 150*fraction_drought
mojave_use_d <- 140*fraction_drought
antelope_use_d <- 220*fraction_drought
ventura_use_d <- 245*fraction_drought
# But cost is a function of abatement, so we need the abatement values instead.
kern_abate_cap_d <- 150-kern_use_d
mojave_abate_cap_d <- 140-mojave_use_d
antelope_abate_cap_d <- 220-antelope_use_d
ventura_abate_cap_d <- 245-ventura_use_d
# DON'T KNOW ABOUT STRINGENCY HERE
# Cost, plug in REDUCTION to each district's cost curve, then take the area under the marginal cost curve for each at the given price/quantity reduction
kern_price_cap_d <- 2.29*kern_abate_cap_d
kern_cost_cap_d <- 0.5*kern_price_cap_d*kern_abate_cap_d
mojave_price_cap_d <- 3.8*mojave_abate_cap_d
mojave_cost_cap_d <- 0.5*mojave_price_cap_d*mojave_abate_cap_d
antelope_price_cap_d <- 2.86*antelope_abate_cap_d
antelope_cost_cap_d <- 0.5*antelope_price_cap_d*antelope_abate_cap_d
ventura_price_cap_d <- 1.78*ventura_abate_cap_d
ventura_cost_cap_d <- 0.5*ventura_price_cap_d*ventura_abate_cap_d
total_cost_cap_d = ventura_cost_cap_d + kern_cost_cap_d + antelope_cost_cap_d + mojave_cost_cap_d
```


*(1)* Total cost: $`r round(total_cost_cap_d,0)`.

*(2)* Individual district costs:

- Kern : $`r round(kern_cost_cap_d,0)`
- Mojave: $`r round(mojave_cost_cap_d,0)`
- Antelope: $`r round(antelope_cost_cap_d,0)`
- Ventura: $`r round(ventura_cost_cap_d,0)`

Antelope County has the highest individual cost with the cap without trade policy in the drought scenario. 


#### Drought and tax on water use

```{r}
# Tax applied to reach 300 AF level, 455 AF reduction 
# Calculate aggregate demand to get the price at which cumulative water use is \300 AF
# Same aggregate cost slope as above
tax_level_d <- 455/aggregate_cost_slope
# Tax of $282.38/AF is applied across all counties. 
# Now we need to see cost to each district, plug the price into the individual MC curves then take the area under each and sum for total. 
# Kern
# tax_level = 2.29A
kern_abate_tax_d <- tax_level_d/2.29
# 123.3 AF
kern_cost_tax_d <- 0.5*tax_level_d*kern_abate_tax_d
# Mojave
# tax_level = 3.8A
mojave_abate_tax_d <- tax_level_d/3.8
#74.31 AF
mojave_cost_tax_d <- 0.5*tax_level_d*mojave_abate_tax_d
#Antelope
# tax_level = 2.86A
antelope_abate_tax_d = tax_level_d/2.86
# 98.74 AF
antelope_cost_tax_d <- 0.5*tax_level_d*antelope_abate_tax_d
# Ventura
# tax_level = 1.78A 
ventura_abate_tax_d = tax_level_d/1.78
# 158.64 AF
ventura_cost_tax_d <- 0.5*tax_level_d*ventura_abate_tax_d
#ventura_abate_tax_d+mojave_abate_tax_d+antelope_abate_tax_d+kern_abate_tax_d
# Total abatement 455 
total_cost_tax_d = 0.5*455*tax_level_d 
# Revenue is the tax times the AF used. 
tax_revenue_d <- tax_level_d*300
```

*(1)* Total cost: $`r round(total_cost_tax_d,0)`.

*(2)* Individual district costs:

- Kern : $`r round(kern_cost_tax_d,0)`
- Mojave: $`r round(mojave_cost_tax_d,0)`
- Antelope: $`r round(antelope_cost_tax_d,0)`
- Ventura: $`r round(ventura_cost_tax_d,0)` 

*(3)* Tax revenue generated by this policy (based on 300 Acre Feet of total use) is $`r round(tax_revenue_d,0)` . This is moderately higher than the tax revenue calculated earlier for the normal (non-drought) scenario. This policy drastically decreases Kern's water consumption so it would not be favored by Kern County. 


#### Drought and cap and trade
From part (a), the cost to each district for the cap without trade was calculated to be: 

- Kern : $`r round(kern_cost_cap_d,0)`, reduction of 90.45 AF. Kern's total post-drought use is `r kern_use_d` AF. 
- Mojave: $`r round(mojave_cost_cap_d,0)`, reduction of 84.42 AF. Mojave's total post-drought use: `r mojave_use_d` AF. 
- Antelope: $`r round(antelope_cost_cap_d,0)`, reduction of 132.66 AF. Antelope's total post-drought use: `r antelope_use_d` AF. 
- Ventura: $`r round(ventura_cost_cap_d,0)`, reduction of 147.74 AF. Ventura's total post-drought use: `r ventura_use_d` AF. 

the total cost: $`r round(total_cost_cap_d,0)`.

```{r}
# Still need to abate 455 AF total, but Mojave and Kern will likely face higher relative MC while Antelope and Ventura will face slightly lower MC. 
trade_mc_d <- tax_level_d
#Kern Q with trade
kern_abate_trade_d <- trade_mc_d/2.29
#kern_abate_cap_d - kern_abate_trade_d
#32.86 increase in abatement
#Mojave Q with trade
mojave_abate_trade_d <- trade_mc_d/3.8
#mojave_abate_cap_d-mojave_abate_trade_d
#-10.11 decrease in abatement 
#Antelope Q with trade
antelope_abate_trade_d <- trade_mc_d/2.86
#antelope_abate_cap_d-antelope_abate_trade_d
#-33.92 decrease in abatement 
#Ventura Q with trade
ventura_abate_trade_d <- trade_mc_d/1.78
ventura_abate_cap_d-ventura_abate_trade_d
#10.91 increase in abatement
# Same allocation/cost as efficient tax, different revenue implications. MC for abatement is the same for each district at these allocations. 
total_cost_trade_d<- 0.5*trade_mc_d*455
```

*(1)* Total cost of this approach is $`r round(total_cost_trade_d,0)` (same as tax).
This is $`r round(total_cost_cap_d - total_cost_trade_d, 0)` lower than the cost of the cap with no trade.

*(2)* Individual district costs: 

- Kern : $`r round(kern_cost_tax_d - trade_mc_d*32.86,0)`

- Mojave: $`r round(mojave_cost_tax_d - trade_mc_d*-10.11,0)`

- Antelope: $`r round(antelope_cost_tax_d - trade_mc_d*-33.92,0)`

- Ventura: $`r round(ventura_cost_tax_d - trade_mc_d*10.91,0)` 

Based on these cost calculations, no districts will prefer cap without trade. Kern and Ventura will prefer cap and trade. Antelope and Mojave will prefer the tax policy. 

*(3)* No tax revenue is created by this policy. 

